---
title: "Code a Valentine!"
author: "International Love Data Week"
date: "Feb 11, 2025"
format: pdf
editor: visual
execute: 
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---

This document contains instructions for creating images for a valentine or art using `R` code. We have lots of helper here to walk you through how to run and modify the code. Please raise your hand if you need help!

You should experiment and get creative with all of the options before deciding which one to include on your Valentine's card. Each option includes an example and then a place for you to modify the code and create something that is unique to you!

# But First: How to Run and Modify Code!

The following gray box is called an `R` **chunk**.

-   To run all of the code in the gray box, click on the green triangle.

-   To run a single line of code, put the cursor on the line you want to run and hit Command+Enter (for a Mac) and Control+Enter (for Windows).

```{r}
2 + 2
5 - 2
```

## One more thing!

We want the plots to show up in the "Plots" window on the right. To make sure that happens, click on the white cog next to the "Render" button and check "Chunk Output in Console."

# Option 1: Heart of Hearts!

## Example

Run the following code to create a heart of hearts.

```{r}
# Load packages
library(tidyverse)
library(emojifont)

# Ensure the exported images are the same size in the Plots window
library(showtext)
showtext_opts(dpi = 1000)

# Add more font options
font_add_google("Merriweather")
font_add_google("Lora")

# Generate dataset with heart coordinates
heart_x <- function(angle) {
  x <- (16 * sin(angle) ^ 3) / 17
  return(x - mean(x))
}

heart_y <- function(angle) {
  y <- (13 * cos(angle) - 5 * cos(2 * angle) - 2 * cos(3 * angle) -
          cos(4 * angle)) / 17
  return(y - mean(y))
}

heart_data <- tibble(
  angle = seq(0, 2 * pi, length.out = 50),
  x = heart_x(angle),
  y = heart_y(angle)
) %>%
  mutate(heart = fontawesome("fa-heart"))



#########
# Experiment with the following items
#########

# Try changing the colors
color_heart = sample(c("deeppink", "purple3", "red3"), 
                     size = 50, replace = TRUE) 
color_text = "deeppink3"
# Add your valentine's name
valentines_name = "Dad!"

# Play around with the size of the hearts
heart_size = 18

# Make the hearts more or less opaque
how_opaque = 0.6

# Font
font_writing = "Lora"

# Font size
font_size = 15

#########

# Graph the hearts and name of valentine
ggplot(data = heart_data, 
       mapping = aes(x = x, y = y, label = heart)) +
  geom_text(alpha = how_opaque,
            color = color_heart,
            size = heart_size,
            family = 'fontawesome-webfont') +
  theme_void() + 
  annotate("text", x = 0, y = 0, 
           label = "Happy Valentine's Day,",
           size = font_size,
           color = color_text,
           family = font_writing) + 
  annotate("text", x = 0, y = -0.25,
           label = valentines_name,
           size = font_size,
           color = color_text,
           family = font_writing) 

ggsave("option1_example.png", dpi = 500)
```

## Your turn!

**Tip**: This website is a good place to find hex codes for colors: <https://coolors.co/palettes/trending>

```{r}

#########
# Experiment with the following items
#########

# Try changing the colors
color_heart = "red3"
color_text = "deeppink3"

# Add your valentine's name
valentines_name = "Insert Name!"

# Play around with the size of the hearts
heart_size = 18

# Make the hearts more or less opaque
how_opaque = 0.6

# Change the font
# Add more font options from Google Fonts
font_add_google("Lora")
font_writing = "Lora"

# Font size
font_size = 15

#########

# Graph the hearts and name of valentine
ggplot(data = heart_data, 
       mapping = aes(x = x, y = y, label = heart)) +
  geom_text(alpha = how_opaque,
            color = color_heart,
            size = heart_size,
            family = 'fontawesome-webfont') +
  theme_void() + 
  annotate("text", x = 0, y = 0, 
           label = "Happy Valentine's Day,",
           size = font_size,
           color = color_text,
           family = font_writing) + 
  annotate("text", x = 0, y = -0.25,
           label = valentines_name,
           size = font_size,
           color = color_text,
           family = font_writing) 

ggsave("option1_example.png", dpi = 1000, width = 7, height = 5,
       units = "in")
```

# Option 2: Many Hearts: Two Ways

## Example

Run the following code to create a wall of hearts, two ways.

```{r}
# Create grid for hearts
hearts <- expand.grid(x = 1:14, y = 1:10) %>%
  mutate(heart = fontawesome("fa-heart")) %>%
  mutate(special = if_else(x == 10 & y == 4, "yes", "no")) %>%
  mutate(cycle = rep(letters[1:5], 28))


# Graph the hearts with a fun pattern
ggplot(data = hearts, 
       mapping = aes(x = x, y = y,
                     label = heart,
                     color = cycle)) +
  geom_text(size = 10,
            family = 'fontawesome-webfont') +
  scale_color_manual(values = c("#277da1", "#f3722c",
                                "#4d908e", "#f9c74f", 
                                "#f94144")) + 
  guides(color = "none") +
  theme_void()

ggsave("option2_example.png", dpi = 1000, width = 7, height = 5,
       units = "in")

```

```{r}
# Graph the hearts with one special heart
ggplot(data = hearts, 
       mapping = aes(x = x, y = y,
                     label = heart,
                     color = special)) +
  geom_text(size = 10,
            family = 'fontawesome-webfont') +
  scale_color_manual(values = c("#277da1", "#f3722c")) + 
  guides(color = "none") +
  theme_void()

ggsave("option3_example.png", dpi = 1000, width = 7, height = 5,
       units = "in")
```

## Your turn!

```{r}

###########
# Pick 5 new colors
colors = c("#277da1", "#f3722c", "#4d908e", "#f9c74f", "#f94144")
###########


# Graph the hearts with a fun pattern
ggplot(data = hearts, 
       mapping = aes(x = x, y = y,
                     label = heart,
                     color = cycle)) +
  geom_text(size = 10,
            family = 'fontawesome-webfont') +
  scale_color_manual(values = colors) + 
  guides(color = "none") +
  theme_void()

ggsave("option2_example.png", dpi = 1000, width = 7, height = 5,
       units = "in")

```

```{r}
###########
# Pick 2 new colors
colors = c("#277da1", "#f3722c")
###########


# Graph the hearts with one special heart
ggplot(data = hearts, 
       mapping = aes(x = x, y = y,
                     label = heart,
                     color = special)) +
  geom_text(size = 10,
            family = 'fontawesome-webfont') +
  scale_color_manual(values = colors) + 
  guides(color = "none") +
  theme_void()

ggsave("option3_example.png", dpi = 1000, width = 7, height = 5,
       units = "in")
```

# Option 3: Grow a Forest

The next two options come from [Danielle Navarro](https://djnavarro.net/), a data scientist and generative artist. She licenses all her code for creating generative art in `R` using creative common licenses so that others can use her code to create their own art!

## Example

Run the following code to grow your own forest.

```{r}
# Load Danielle's package
library(flametree)

# Pick some colors
shades = c("#5F0F40", "#9A031E", "#E36414", "#FB8B24")

# Grow the trees
dat = flametree_grow(seed = 8,
                      time = 6, 
                      trees = 5)

# Draw the trees
dat %>% 
  flametree_plot(
    background = "#0F4C5C",
    palette = shades
  ) +
  annotate("text", x = -2.5, y = 1, 
           label = "Hug\n a\n Tree",
           size = 12, color = "#FB8B24",
           family = font_writing)

ggsave("option4_example.png", dpi = 500)
```

## Your turn!

```{r}
#########
# Change the following
#########

# Determine how many trees you want
number_trees = 5

# Pick a random number for growing the trees 
# Impacts their location
number_random = 8

# Determine how much growing (larger number = more growth)
number_growing = 6

# Pick some new colors
shades = c("#5F0F40", "#9A031E", "#E36414", "#FB8B24")

# Pick a background color
color_background = "#0F4C5C"

# Add a message
# Tips: "" = no message
# \n = new line
message = "Hug\n a\n Tree"

# Pick the size of the message
message_size = 12

# Pick the x location and y location of the message
message_x = -2.5
message_y = 1

# Pick the color of the message
message_color = "#FB8B24"

# Change the font
# Add more font options from Google Fonts
font_add_google("Lora")
message_font = "Lora"

#########

# Grow the trees
dat <- flametree_grow(seed = number_random,
                      time = number_growing, 
                      trees = number_trees)

# draw the plot
dat %>% 
  flametree_plot(
    background = background_color,
    palette = shades
  ) +
  annotate("text", 
           label = message,
           size = message_size,
           x = message_x, y = message_y, 
           color = message_color,
           family = message_font)

ggsave("option4_example.png", dpi = 500)
```

# Testing

<https://blog.djnavarro.net/posts/2021-10-19_rtistry-posts/>

```         
```

https://jasmines.djnavarro.net/

```{r}
library(jasmines)
use_seed(4) %>%
  scene_discs(
    rings = 4, points = 5000, size = 6
  ) %>%
  mutate(ind = 1:n()) %>%
  unfold_warp(
    iterations = 1,
    scale = .5, 
    output = "layer" 
  ) %>%
  unfold_tempest(
    iterations = 20,
    scale = .011
  ) %>%
  style_ribbon(
    palette = "magma",
    colour = "ind",
    alpha = c(.5,.1),
    background = "gray90"
  )

ggsave("option4_example.png", dpi = 500)
```

```{r}
library(jasmines)
use_seed(10) %>%
  scene_discs(
    rings = 3, points = 5000, size = 5
  ) %>%
  mutate(ind = 1:n()) %>%
  unfold_warp(
    iterations = 1,
    scale = .5, 
    output = "layer" 
  ) %>%
  unfold_tempest(
    iterations = 20,
    scale = .01
  ) %>%
  style_ribbon(
    palette = palette_named("vik"),
    colour = "ind",
    alpha = c(.1,.1),
    background = "oldlace"
  )
```
